An FPGA Implementation of a Long Short-Term Memory Neural Network

### 2016-07-05 paper/paper.tex
Added the IEEE paper for Reconfig

\chapter{Problem Characterization}\label{chap:theorBack}

\section{Theoretical Background}\label{sec:tb}

\subsection{Basic Concepts of Machine Learning}\label{sec:tb_basicCon}
Machine Learning is a field of Computer Science that studies the development of mathematical techniques that allow software to learn autonomously, without an explicit description of each rule of operation. Its goal is to extract latent features from the data that allow an immediate classification of each input data into a particular class -- the catch is that there is no previous rule formulation, but instead we have an adaptive model that adjusts is parameters according to the input data it receives, improving the estimates it yields as it receives new input samples.

Let us consider a practical example. For instance, suppose we want to build a program that given an input audio waveform representation of a spoken word, it matches it into a particular word of a dictionary. We could, of course, devise a set of rules and exceptions for each word analysing some of its features (perhaps the Fourier representation of each one, and, from it, manually finding the appropriate rules for each), but apart from being a very complex task, it wouldn't be a scalable solution, given the enormous number of words in each language. The approach taken by Machine Learning is different, and instead of manually processing each waveform, we build a large dataset, of size $N$, containing the waveforms of several words $\left[ \mb{x}_1 ,\; \mb{x}_2 ,\;\mb{x}_2 ,\; \dots ,\;\mb{x}_N \right]$ -- we call this dataset the \textbf{training dataset} -- and we feed it to our model. Each of the $i$-th data point was previously labelled, and in fact we feed each training data point $\mb{x}_i$ \emph{along} with its corresponding label $t_i$, so that the model can adapt its parameters accordingly to the \emph{target value} it is supposed to classify. This set of labels $\mb{t} = \left[ t_1 ,\; t_2 ,\; t_3 ,\; \dots ,\; t_N \right]$ is called the \textbf{target vector}.

We are, then, left with the following question: how can the model quantitatively evaluate the quality of its current set of parameters? That could be achieved in a number of ways, but the most usual is using a \textbf{Cost Function} that, as the name suggests, measures the cost of each wrong classification of the model. The model then evolves in a way that minimizes the cost function. A usual choice for the cost function is the \textbf{sum of squares error}, given the Gaussian Noise assumption. Mathematically, if $y_{\theta}^i = y_{\theta}(x_i)$ is the prediction for the input data point $x_i$ with label $t_i$, given the current set of parameters $\theta$, the cost function using this metric is given by

\begin{equation}\label{eq:costfunctionFund}
	J(\theta) = \frac{1}{2} \sum_{i=1}^N \left( y_{\theta}^i - t_i \right)^2.
\end{equation}

Sometimes, instead of applying the raw data to the model, we can apply some sort of \emph{preprocessing} to the data to extract the relevant features from it. For instance, instead of just feeding a raw image, we can perform several operations like edge detection or low-pass filtering, and apply them in parallel. In cases of highly dimensional data (i.e. each data vector has a very high number of features), we can apply techniques like \textbf{Principal Component Analysis} to reduce the feature space to a smaller dimension one, where the previous features were combined into two or three new features that pose themselves as the most relevant.

The problems described above are, in fact, a subset of the problems that Machine Learning tries to address. These problems are called \textbf{classification problems}, because for each input data point, our model tries to fit it into the most appropriate class. But we can also address \textbf{regression problems} where the output is not limited to a discrete set of values but rather a continuous interval. On the other hand, the Neural Network that this work will implement addresses a special kind of classification problem, where the classification decision is influenced not only by the current input sample, but also by a given \emph{window of samples} that trail the current sample.

In summary, the most typical setting for a Machine Learning problem is having a large \textit{input dataset} which we use to \textit{train} our model (i.e. allowing him to dynamically adapt its set of parameters $\theta$), in order to produce an output label $y_i$ for each of them that minimizes a quality metric, the \textit{Cost Funcion}, which can be chosen to the sum of squared differences, the log-loss, or any other appropriate  mathematical relationship  between the estimate $y_i$ and the correct label $t_i$.

Now that the basic Machine Learning concepts have been presented, I will discuss, henceforth, one of the most important algorithms that address the supervised classification problems, the \emph{Artificial Neural Networks} that will be discussed in Section~\ref{sec:theorBack_ann} as a contextual introduction to the main theme of the thesis, which will be Recurrent Neural Networks (Section~\ref{sec:theorBack_rnn}), namely the \textbf{Long Short-Term Memory Networks} (Section~\ref{sec:theorBack_lstm}), both of which are improvements over the initial formulation of the ANNs. These two last networks branch even further from these set of problems, and are usually employed in \textit{Deep Learning} tasks, where we try to extract even higher level information from data at the expense of increased model complexity.

\subsection{Artificial Neural Networks}\label{sec:theorBack_ann}

Artificial Neural Networks (ANN) are mathematical structures that, as the name suggests, try to mimic the basic way of how a human brain works. ANN's building blocks, like their biological counterpart, model the high-level behaviour of biological neurons, in the sense that they neglect unnecessary biological aspects (such as modeling all the voltages across the neuron and all its electromagnetic interactions), and only retain its fundamental underlying mathematical function, which is a weighted linear combination of its inputs subject to a \emph{activation function}, i.e. a function that outputs a decision value depending on its inputs.  Mathematically, we have

\begin{equation}
	y = f(\mb{w}^T \mb{x})
\end{equation}
where $\mb{w} = \left[ w_0 \; w_1 \; w_2 \;  w_3 \;  \dots \; w_n \right]$ is the input weight vector, $\mb{x} = \left[ x_0 \; x_1 \; x_2 \; x_3 \; \dots \; x_n \right]$ the input vector, $b_0$ is the bias factor and $f(\cdot)$ is the chosen activation function. Furthermore, we call the scalar quantity $a = \mb{w}^T \mb{x}$ the \textbf{activation}, since its value determines how the activation function will behave. Figure~\ref{fig:modelNeuron} exemplifies the roles of these variables within our neuron model, and compares each part of it with the biological counterpart.

\begin{figure}[H]
	\begin{subfigure}{0.5\linewidth}
		\centering
		\includegraphics[width=0.9\linewidth]{figures/neuron.eps}
		\caption{Neural Network Node}
		\label{fig:modelNeuron_a}
	\end{subfigure}
	\begin{subfigure}{0.5\linewidth}
		\centering
		\includegraphics[width=0.9\linewidth]{figures/MultipolarNeuron.png}
		\caption{Biological Neuron Diagram}
		\label{fig:modelNeuron_b}
	\end{subfigure}

    \caption[Representations of a Neuron in an ANN~\ref{fig:modelNeuron_a} and a biological multi polar neuron~\ref{fig:modelNeuron_b}]{In Figure~\ref{fig:modelNeuron_a}. each input feature $x_i$ is weighted by its corresponding weight $w_i$. During the training procedure, these weights are adjusted so that the output $y$ approaches the target value. In Figure~\ref{fig:modelNeuron_b}, we see the diagram of an actual multi polar neuron. The dendrites, where the stimuli are received, plays a role similar to that of the input nodes. The axon transmits the signal to the synaptic terminals, that are similar to the $y$ output. Source: \href{https://en.wikipedia.org/wiki/Neuron\#/media/File:Neuron\_Hand-tuned.sv://en.wikipedia.org/wiki/Neuron\#/media/File:Blausen\_0657\_MultipolarNeuron.png}{Wikipedia}}
	\label{fig:modelNeuron}
\end{figure}

As far as the activation function is concerned, we can have several types. An immediate choice would be the \textbf{Binary Step Function} that  outputs -1 if the activation is \textbf{below} a given threshold and 1 otherwise. There can also be \textbf{real valued activation functions}, whose output is not binary, but rather that of a continuously differentiable function, such as the logistic sigmoid $\sigma(a) = \frac{1}{1 + \mathrm{e}^{-a}}$ or the hyperbolic tangent $\tanh(a)$. This aspect will prove useful for the usual training methods, that involve the computation of derivatives. In Figure~\ref{fig:activFunc} these activation functions are plotted.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\linewidth]{figures/activFunc.eps}
    \caption[Possible activation functions of choice for a neuron]{Three different activation functions. As you can see, the hyperbolic tangent has the same extreme values as the sign step function, but has a smooth transition between them, which can be interpreted as a \emph{soft decision} in the more ambiguous middle region, reflecting the degree of uncertainty on the decision. On the other hand, the sigmoid function goes from zero to one, and is also smooth like the hyperbolic tangent.}
	\label{fig:activFunc}
\end{figure}

A neuron by itself can be thought of as a simple linear regression, where we optimize the weight of each feature according to a target value, or function. While important in some applications, the main interest in ANN is to evaluate increasingly more complex models, and not a simple linear regression. This is achieved by \emph{chaining} nodes to one another, connecting the output of a given node to one of the inputs of another. We call \emph{layers} to a group of these nodes that occupy the same hierarchical position. There can be any number of layers, with any number of nodes, but most implementations generally have 3 layers: the \emph{input} layer, the \emph{hidden} layer (in the middle) and the \emph{output} layer. Figure~\ref{fig:neuralnet} suggests a possible structure for a 3 layer ANN.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\linewidth]{figures/ann.eps}
    \caption[Three layer ANN]{A three layer ANN. We have omitted some of the connections in the hidden layer, for simplification purposes. $\mb{w}_1$ represents the weight matrix of the input layer, $\mb{w}_2$ the weight matrix of the connections between the input layer and the hidden layer, and $\mb{w}_3$ the weight connections between the hidden and the output layer. $f_{ij}(\dots)$ is the activation function of the $j$-th neuron of the $i$-th layer. Since they can be different, I chose different indexes to each.}
	\label{fig:neuralnet}
\end{figure}

Regarding the training of ANNs, it is performed through a two-step process: first, a \textbf{feed-forward} step where the input is applied, and the activations are evaluated in succession up to the output neurons; then, we perform the \textbf{backpropagation} step, where we calculate the errors in each of the nodes (the so-called \textbf{deltas} of equation~\ref{eq:deltaj}), but now from the output to the input: the weights are updated and optimized using an iterative  method called \textit{Gradient Descent}, where if $\tau$ is the current time step, the next update on the weight matrix $\mb{w}^{(\tau+1)}$ is given by

\begin{equation}\label{eq:gradDesc}
	\mb{w}^{(\tau+1)} = \mb{w}^{(\tau)} - \eta \nabla E(\mb{w}^{(\tau)})
\end{equation}
where $E(\cdot)$ is the error function. As we can see, the weight matrix is moved in the direction that minimizes the error function the most, and $\eta$ controls how fast this is achieved, being the reason why it is called the \textbf{learning rate}.

The computation of gradient of the error function comprises the evaluation of its derivatives with respect to each weight of all network connections, $w_{ij}$. They are

\begin{equation}\label{eq:partialE}
	\frac{\partial E}{\partial w_{ji}} = \delta_j f(a_i)
\end{equation}
where $f(\cdot)$ is the activation function of the neuron and

\begin{equation}\label{eq:deltaj}
	\delta_j = f'(a_j) \sum_k w_{kj} \delta_k .
\end{equation}
The interpretation of these equations is simple. If $w_{ji}$ is the weight of the connection between the neuron $j$ we are considering and a neuron $i$ in a previous layer, then the sum over $k$ relates to all the neurons in the \emph{next} layer to which $j$ connects: this way, since the update of $w_{ji}$, according to~\ref{eq:gradDesc}, is given by

\begin{equation}\label{eq:update}
	w_{ji}^{(\tau+1)} = w_{ji}^{(\tau)} - \eta \frac{\partial E}{\partial w_{ji}}
\end{equation}
we see that, from~\ref{eq:partialE}  it simply is the product of the error of the current neuron, $\delta_j$, with the output of the previous neuron $f(a_i)$. In turn, from~\ref{eq:deltaj}, we see the recurrence relationship between it and the weighted sum of all posterior neurons that connect to it. Hence, the name backpropagation is now clear: we are, in fact, propagating the errors backwards into the neuron of interest, weighted by the corresponding weight, but now \emph{backwards} instead of forward, as before. For the output units, the $\delta_j$ is simply the difference between the produced output and the corresponding label for that sample. This two-step process is performed for every data point in out dataset. %%Place where I can include an algorithmic overview?
For a complete proof of the above formulas, see~\cite[chap. 5.3.1]{Bishop2006}.

\subsection{Recurrent Neural Networks}\label{sec:theorBack_rnn}
A Recurrent Neural Network (RNN) is, essentially, a regular ANN where some neurons (especially in the hidden layer) have \emph{feedback connections} to themselves, i.e. their outputs are fed as inputs. The relevance of this different structure is the possibility to retain \emph{sequence} information about the data. Before, each incoming data point only contributed to the training of the network, but the information about the correlation between themselves and the data points that preceded them did not influence the training step. Temporal relationship is disregarded and each data point considered conditionally independent of any other. This is obviously not necessarily true, and in fact there are many cases where the correlation between data points is high for those closely spaced in time, such as in video signals, audio signals, or other kinds of \emph{temporal sequences} of data. Therefore, the feedback connection of the neuron to himself acts as a kind of \emph{memory element} that takes into account in the present decision, the history of decisions previously taken, and hence the previous data.

Figure~\ref{fig:recurrneuron} suggests a possible structure for a neuron of a hidden layer in an RNN, and also an alternate representation, where the structure is unfold through time.

\begin{figure}[H]
	\begin{subfigure}{0.5\linewidth}
		\centering
		\includegraphics[width=0.34\linewidth]{figures/recneuron.eps}
		\caption{Recurrent Neuron}
		\label{fig:recneuron}
	\end{subfigure}
	\begin{subfigure}{0.5\linewidth}
		\centering
		\includegraphics[width=0.9\linewidth]{figures/recneuron_unf.eps}
		\caption{Recurrent Neuron, unfolded through time}
		\label{fig:recneuron_unf}
	\end{subfigure}

    \caption[Representation of a Recurrent Neuron (Figure~\ref{fig:recneuron}) and a Recurrent Neuron unfolded through time (Figure~\ref{fig:recneuron_unf})]{In Figure~\ref{fig:recneuron}, a recurrent neuron is depicted, where the output is fed back to the input, weighted by $\theta_h$. Figure~\ref{fig:recneuron_unf}, an unfolding through time of that same neuron is performed. The basic idea is that feeding the output back to the input is similar to feeding the output to the input of the neuron \emph{at a previous time step}: this way, we linearize the structure. Considering a training epoch of $T$ samples, corresponds to having $T$ unfolded neurons/layers}
	\label{fig:recurrneuron}
\end{figure}

The training of an RNN is usually performed using a variation of Backpropagation, called \textbf{Backpropagation Through Time} (BPTT), that as the name implies, performs the same backpropagation procedure discussed for the ANNs, but now taking into account the unfolding of the network through a fixed training epoch $T$ like Figure~\ref{fig:recneuron_unf}. Due to this very fact, this training procedure is memory and performance consuming, and so it will not be used in my final work, but instead a novel approach, the \textbf{Simultaneous Perturbation Stochastic Approximation} will be evaluated.

Even though RNNs outperform static ANNs in sequence recognition problems~\cite{Bengio1991}, they fail to retain long-term dependencies. Of course that the weight training process is itself a form of memory, but the problem is that the weight update is much slower than the activations~\cite{Yoshua01}, and therefore this memory only retains short-term dependencies. This is because of the so-called \textbf{Vanishing Gradients Problem}~\cite{Yoshua94,Yoshua01}, where the error decays exponentially through time, and the impact of previous incoming data points on the training of the weights, and thus the current decision, quickly decreases.

\subsection{Long Short-Term Memory Networks}\label{sec:theorBack_lstm}
To overcome the issue of failing to remember long-term dependencies, Hochreiter and Schmidhüber proposed, in 1997, a novel approach to the RNNs called the \textit{Long-Short Term Network}~\cite{Hoch97}. This section explains the main idea of this approach (Section ~\ref{sec:struct_lstm}), and also how it is trained (Section~\ref{sec:training_lstm}), serving as a support for the work of this thesis.
Although originally formulated in 1997, its formulation has been incrementally updated in~\cite{Gers00} and~\cite{Gers2000}, and the most current version is the one in~\cite{Graves05}. One of the inital proposers of LSTM, Prof. Jürgen Schmidhüber, did a survey on the most common variations of the model last year~\cite{Greff15}, and this will the basis of this short theoretical presentation, as well as the work that will be developed in this thesis.

\subsubsection{Structure, Operation and Equations}\label{sec:struct_lstm}
A single LSTM neuron is presented in Figure~\ref{fig:lstmneuron}. As we can see from the picture, we still have the recurrent connections from the regular RNNs, but now there are multiple entry points that control the flow of information through the network. Although omitted from the picture, all the gates are biased, as is suggested in Equations~\ref{eq:equationsLSTM}. The main components, their role and relevance, are explained as follows

\begin{figure}[H]
	\centering
	\includegraphics[width=0.9\linewidth]{figures/lstmneuron.png}
    \caption{A complete LSTM neuron, with all the features as described in~\cite{Graves05}. Source:~\cite{Greff15}}
	\label{fig:lstmneuron}
\end{figure}

@@

\begin{align}
    \mb{z}^{(t)} & = g(\mb{W}_z \mb{x}^{(t)} + \mb{R}_z \mb{y}^{(t-1)} + \mb{b}_z) \notag\\
    \mb{i}^{(t)} & = \sigma(\mb{W}_i \mb{x}^{(t)} + \mb{R}_i \mb{y}^{(t-1)} + \mb{p}_i \odot \mb{c}^{(t-1)} + \mb{b}_i) \notag\\
    \mb{f}^{(t)} & = \sigma(\mb{W}_f \mb{x}^{(t)} + \mb{R}_f \mb{y}^{(t-1)} + \mb{p}_f \odot \mb{c}^{(t-1)} + \mb{b}_f) \notag\\
    \mb{o}^{(t)} & = \sigma(\mb{W}_o \mb{x}^{(t)} + \mb{R}_o \mb{y}^{(t-1)} + \mb{p}_o \odot \mb{c}^{(t)} + \mb{b}_o) \notag\\
    \mb{c}^{(t)} & = \mb{i}^{(t)} \odot \mb{z}^{(t)} + \mb{f}^{(t)} \odot \mb{c}^{(t-1)} \notag \\
    \mb{y}^{(t)} & = \mb{o}^{(t)} \odot h(\mb{z}^{(t)}) \label{eq:equationsLSTM}
\end{align}
where $\odot$ is the Hadamard multiplication. The $i$-th element of the previous vectors in bold corresponds to the value of the gate of the $i$-th neuron of the layer, which is a very convenient and compact representation of the whole layer. Furthermore, if the layer has $N$ LSTM neurons and $M$ inputs (i.e. the size of the layer that precedes this), we see that the input weight matrices $\mb{W}_*$ and the  recurrent weight matrices $\mb{R}_*$ all have size $N \times M$, and that the bias weight matrices $\mb{b}_*$, the peephole weight matrices $\mb{p}_*$ and the matrices $\mb{z}^{(t)}$ through $\mb{c}^{(t)}$ have size $N \times 1$. Although useful for a high-level description in a programming language, the matrix representation  may not be suitable for a direct HDL port. In that case, in order to represent the operation of the $i$-\textit{single neuron}, all the above equations still hold, but instead of \textit{vectors} of outputs, we will have a single scalar output, and we only use the appropriate row in the weight matrices $\mb{W}_{*}$, $\mb{R}_{*}$ and the remaining.

\subsubsection{Training -- SPSA}\label{sec:training_lstm}
Since this work aims to perform on-chip learning, it is important to find a suitable learning scheme. Since I am aiming at an hardware implementation, and although the memory resources of current FPGAs are abundant, one must find an algorithm that uses the less memory as possible (at the smallest performance cost possible), in order to use the additional memory, for instance, to add additional LSTM cells that can improve the performance of our system. 
That being said, we see that the usual training algorithms for LSTM cells -- i.e. Real-Time Recurrent Learning (RTRL), Truncated Backpropagation Through Time (BPTT) or a mixture of both~\cite{Graves05}~\cite{Hoch97}~\cite{Greff15} -- usually involve the storage of the deltas of each layer for every time instant in the training epoch (from $0$ to $T$), which is a highly non-scalable solution both in terms of memory and performance. A most efficient approach to training times series dependant structures like LSTM is the use of \textbf{Simultaneous Perturbation Stochastic Approximation} (SPSA)~\cite{Spall98}.  The main idea of this technique is, instead of explicitly evaluating the gradients for the cost function at each time step, to perform a random walk in the neighbourhood of the current weight matrix, in the weight space, and approximate the new weight update by the approximation of the gradient of the cost function that resulted from that random walk. The weight update for the $i$-th weight at the time step $t$ is given by

\begin{equation}\label{eq:spsa_weightUpdate}
    \Delta w_i^{(t)} = \frac{J(\mb{w}^{(t)} + \beta \mb{s}^{(t)}) - J(\mb{w}^{(t)})}{\beta s_i^{(t)}}
\end{equation}
where $\beta$ is the magnitude of the perturbation to be introduced, $\mb{s}^{(t)}$ is a \textit{sign vector} whose $i$-th element, $s_i^{(t)}$, is either $-1$ or $1$. This way, we see that every weight is \textbf{randomly} incremented either by $-\beta$ or $\beta$, and we only need to keep a duplicate of the weight matrix with the perturbations, and we only need to evaluate the cost function twice per incoming sample. As for the \textit{update rule}, we have

\begin{equation}\label{eq:spsa_updateRule}
    w_i^{(t+1)} = w_i^{(t)} - \eta \Delta w_i^{(t)}
\end{equation}
where $\eta$ is the learning rate, and $\Delta w_i^{(t)}$ is the update for the $i$-th weight evaluated in~\ref{eq:spsa_weightUpdate}.
According to the analysis performed in~\cite{Maeda05}, performing a two-term Taylor expansion at $\mb{w} = \mb{w}^{(t)}$, and taking the expected value of the equation, we get that there is a $\mb{w}_{S1}$ so that

\begin{equation}\label{eq:taylor}
    \Delta w_i^{(t)} = s_i^{(t)} (\mb{s}^{(t)})^T \frac{\partial J(\mb{w}^{(t)})}{\partial \mb{w}} +
                        \frac{c s_i^{(t)}}{2} (\mb{s}^{(t)})^T \frac{\partial^2 J(\mb{w}^{(t)})}{\partial \mb{w}^2}  (\mb{s}^{(t)}) 
\end{equation}
If the random sign vector $\mb{s}^{(t)}$ is chosen carefully so that the expected value of the vector at the $i$-th position is zero, $E(s_i^{(t)})$, and the signs of two different components $i$ and $j$, at different time-steps, are independent (i.e. $E(s_i^{(t_1)}s_j^{(t_2)}) = \delta_{ij}\delta_{t_1 t_2}$, where $\delta$ is the Kronecker fuction), taking the \emph{Expected Value} of equation~\ref{eq:taylor} yields

\begin{equation}\label{eq:spsa_gradProof}
    E(\Delta w_i^{(t)}) = \frac{\partial J(\mb{w}^{(t)})}{\partial w_i^{(t)}}
\end{equation}
that is, the \emph{expected value} of the weight update approximates the gradient of the cost function relative to that weight, and so the learning rule is a special form of \textit{Stochastic Gradient Descent}.

One last comment to be made concerning the update rule, is the hypothetical need for \textbf{limits} to the weight values, when the update rule exceeds $|w_{\text{max}}|$ (in that case, we set $w_i^{(t+1)} = \pm w_{\text{max}}$), which sometimes might be needed if the behaviour of the weight update is not appropriate~\cite{Maeda05}.

>>>>

\documentclass{IEEEtran}

%% Additional Packages
\usepackage{cite}
\usepackage{graphicx}
%% Macros
\newcommand{\mb}[1]{\mathbf{#1}}

\title{An FPGA Implementation of a Long Short-Term Memory Neural Network}

%% COMMENT THIS FOR SUBMISSION
\author{\IEEEauthorblockN{Jose Pedro Castro Fonseca} \\ 
\IEEEauthorblockA{Faculty of Engineering University of Porto \\
Porto, Portugal}}
\IEEEspecialpapernotice{(Working Version)}

\begin{document}

\maketitle

\begin{abstract}
This work proposes a hardware architecture for a Long Short-Term Memory (LSTM) Neural Network, aiming
to outperform software implementations, by exploiting the inherent parallelism in it.  
The main design decisions are presented, along with the proposed network architecture. A description of the main
building blocks of the network is also presented. The network is synthesized for various sizes,
and performance results are presented and analysed. The synthesized network achieves a 195 times speed-up
over a custom-built software network, proving the benefits of parallel computation for this kind of network.
\end{abstract}

\section{Introduction}\label{sec:intro}
\IEEEPARstart{N}{eural} Networks are one of the most commonly used techniques in Deep Learning. This
particular type of network, LSTM, is a recursive network, given that the neuron outputs in a certain 
time step are also fed as the input in the next time step, and that way their memory elements can make
sense of patterns within data sequences, unlike classical recursive neural networks.
These algorithms have been profusely implemented in software, and their practical applications are plentiful. 
However, the benefits of the inherent parallelism offered by a dedicated hardware
platform are not availed, and there are relatively few implementations of Machine Learning algorithms in 
these kind of platforms.

Hardware platforms would achieve a considerable speed-up over software implementations, which could prove
useful for high data throughput systems, where the calculation overhead is critical and limits the performance of the systems.
Furthermore, a hardware implementation can even benefit \emph{offline} Deep Learning tasks by providing the results
of a given experiment faster than in a regular CPU, yielding an increase in scientific productivity.

Hitherto, there is only one implementation~\cite{Chang15}, but its performance is undermined
by the external memory access. This implementation aims to make use of internal FPGA memory resources, and therefore
achieve a higher speed performance.

In overview, this article begins by explaining what are LSTM networks -- and their mathematical details -- in Section~\ref{sec:lstmnn}, and by providing a quick overview of their state of the art in Section~\ref{sec:relwork}.
Section~\ref{sec:proparch} outlines the proposed hardware architecture and its main constituent modules, and the performance and synthesis results are reported in Section~\ref{sec:results}. A final concluding remark is given in Section~\ref{sec:concl}.

\section{LSTM Neural Networks}\label{sec:lstmnn}
Although originally formulated in~\cite{Hoch97}, its formulation has been incrementally updated in~\cite{Gers00} and~\cite{Gers2000}, and the most current version is the one in~\cite{Graves05}. One of the inital proposers of LSTM, Prof. Jürgen Schmidhüber, did a survey on the most common variations of the model last year~\cite{Greff15}, and this will the basis of this short theoretical presentation, as well as the work that will be developed in this thesis.

A single LSTM neuron is presented in Figure~\ref{fig:lstmneuron}. As we can see from the picture, we still have the recurrent connections from the regular RNNs, but now there are multiple entry points that control the flow of information through the network. Although omitted from the picture, all the gates are biased, as is suggested in Equations~\ref{eq:equationsLSTM}. The main components, their role and relevance, are explained as follows

\begin{figure}[!t]
	\centering
	\includegraphics[width=2.5in]{figures/lstmneuron.png}
    \caption{A complete LSTM neuron, with all the features as described in~\cite{Graves05}. Source:~\cite{Greff15}}
	\label{fig:lstmneuron}
\end{figure}

@@

\begin{eqnarray}
    \mb{z}^{(t)} & = g(\mb{W}_z \mb{x}^{(t)} + \mb{R}_z \mb{y}^{(t-1)} + \mb{b}_z) \nonumber\\
    \mb{i}^{(t)} & = \sigma(\mb{W}_i \mb{x}^{(t)} + \mb{R}_i \mb{y}^{(t-1)} + \mb{p}_i \odot \mb{c}^{(t-1)} + \mb{b}_i) \nonumber\\
    \mb{f}^{(t)} & = \sigma(\mb{W}_f \mb{x}^{(t)} + \mb{R}_f \mb{y}^{(t-1)} + \mb{p}_f \odot \mb{c}^{(t-1)} + \mb{b}_f) \nonumber\\
    \mb{o}^{(t)} & = \sigma(\mb{W}_o \mb{x}^{(t)} + \mb{R}_o \mb{y}^{(t-1)} + \mb{p}_o \odot \mb{c}^{(t)} + \mb{b}_o) \nonumber\\
    \mb{c}^{(t)} & = \mb{i}^{(t)} \odot \mb{z}^{(t)} + \mb{f}^{(t)} \odot \mb{c}^{(t-1)} \nonumber \\
    \mb{y}^{(t)} & = \mb{o}^{(t)} \odot h(\mb{z}^{(t)}) \label{eq:equationsLSTM}
\end{eqnarray}
where $\odot$ is the Hadamard multiplication. The $i$-th element of the previous vectors in bold corresponds to the value of the gate of the $i$-th neuron of the layer, which is a very convenient and compact representation of the whole layer. Furthermore, if the layer has $N$ LSTM neurons and $M$ inputs (i.e. the size of the layer that precedes this), we see that the input weight matrices $\mb{W}_*$ and the  recurrent weight matrices $\mb{R}_*$ all have size $N \times M$, and that the bias weight matrices $\mb{b}_*$, the peephole weight matrices $\mb{p}_*$ and the matrices $\mb{z}^{(t)}$ through $\mb{c}^{(t)}$ have size $N \times 1$.

\section{Related Work}\label{sec:relwork}

\section{Proposed Architecture}\label{sec:proparch}

\subsection{Activation Function Calculation}

\subsection{Matrix-vector Dot Product Calculation}

\subsection{Weight storage}

\subsection{Network Architecture}

\section{Results}\label{sec:results}

\subsection{Validation}

\subsection{Synthesis}

\subsection{Performance}

\section{Conclusion}\label{sec:concl}

\bibliographystyle{IEEEtran}  
\bibliography{bibliography}

\end{document}
